\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-10-20 icra-2021.tex
updating the algorithm section and some consistency fixes

The algorithm uses the concept to select the adjustments s.t. $\varphi_k(\mathbf{p}_k,\mathbf{c}_k^0)$ has the highest energy value. It guides the UAV to the new position $\mathbf{p}_{k+1}$ using the vector field of $\varPhi:=\varphi_k(\mathbf{p}_k,\mathbf{c}_k^0)$, deriving the direction to follow--the desired velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is the gradient, $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference.

@@

Let us consider a Fourier series $h:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ for the purpose of energy modeling of the mission
\begin{equation}\label{eq:fourier}
  h(t)=\sum_{n=0}^{r}{a_n\cos\left({nt/\xi}\right)+b_n\sin{\left({nt/\xi}\right)}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ the Fourier series coefficients.

@@

The energy cost of the computations is assessed using \powprof{}, an open-source modeling tool presented in our previous work~\cite{seewald2019coarse}, that measures software configurations empirically and builds an energy model. Specifically, the tool builds a linear interpolation, one per each task. It requires the user to implement the mission as a ROS system with one or more ROS nodes changing the computational load by node-specific ROS parameters. A way to simulate the change of the computations.

@@

%The main goal of the algorithm is to find an optimal control action $\mathbf{u}^0$ for the current state $\hat{\mathbf{q}}$--estimated from sensors' measurements--from the optimal control law $\mathbf{u}^0=:\kappa(\hat{\mathbf{q}})$. This is achieved by solving online a finite horizon optimal control problem by the hand of a modification of a model predictive control (MPC) algorithm~\cite{rawlings2017model}.

Given an initial mission $\mathcal{M}_1$ (TEE $\varphi_1$, tasks $\Psi$, computations $\mathbf{s}_1\in\mathbb{S}_1$, and adaptations $\mathbf{c}_1\in\mathbb{C}_1$), the algorithm produce a valid mission evolution.

\begin{defn}[Valid mission evolution]\label{def:valid}
  At a generic time instant $k\in[t_f]$ a mission evolution $\mathcal{M}_{k}$ is valid if its control sequence is valid
  \begin{equation}\begin{split}
    \mathbf{u}^a_{k}=\{(\mathbf{c}_{k},\mathbf{s}_{k})\mid&(\varphi(\mathbf{p}_{k-1},\mathbf{c}_{k-1}),\Psi(\mathbf{s_{k-1}}))\in\mathcal{M}_{k-1}\\
    &\Longrightarrow(\varphi(\mathbf{p}_{k},\mathbf{c}_{k}),\Psi(\mathbf{s}_{k}))\in\mathcal{M}_{k}\}.
  \end{split}\end{equation}
\end{defn}

Let us proof that if the mission evolution is valid, the instantaneous energy consumption is a linear combination of the state from the Equation~(\ref{eq:state-perf}).

\begin{thm}[State output linearity]\label{thm:state-vs-energy}
  Consider the mission from Definition~\ref{def:mission}, the valid mission evolution from~\ref{def:valid}, and assume Assumption~\ref{assm:periodic} holds. Likewise in Lemma~\ref{lem:state-vs-energy}, the model behaves ideally. 
  Then, the instantaneous energy consumption $y_k$ is a linear combination of the state $\mathbf{q}_k$ and $\mathcal{M}_k$ produce a valid mission evolution $\mathcal{M}_{k+1}\,\forall k\in[t_f-1]$.
\end{thm}
\begin{proof}
  *\\
\end{proof}

@@

\begin{center}\textit{Ok, so I need to fix this algorithm properly. But the sketch is there (the position is missing but I'll make a wrapper tomorrow morning)!}\end{center}

\begin{algorithmic}[1]
  \Procedure{}{$\mathcal{M}_1$, $\mathbf{p}_1$, $\mathbf{q}_1$, $K$, $v$, $\varepsilon$}
  \State $k\gets 2$
  \State $\mathbf{z}_{k-1},\hat{\mathbf{q}}_{k-1}\gets \mathbf{q}_1$
  \State $\mathbf{v}_{k-1},\mathbf{u}_{k-1}\gets {\mathbf{u}_{k-1}(\overline{\mathbf{u}}_{k-1}^a,\mathbf{u}_{k-2}^a)}$
  \While{$k \leq t_f$}
    \Do
      \State $\mathbf{v}\gets\argmax_{\mathbf{v}}{\sum_{i=k-1}^{k+N-2}}{l(\mathbf{z}_i,\mathbf{v}_i)+V_f(\mathbf{z}_{N+1})}$
      \State $\mathbf{u}_{k-1}\gets \mathbf{v}+K(\hat{\mathbf{q}}_{k-1}-\mathbf{z}_{k-1})$
      \State $\hat{\mathbf{q}}_k^-\gets A\hat{\mathbf{q}}_{k-1} + B\mathbf{u}_{k-1}$
      \If{$C\hat{\mathbf{q}}_k^-\notin\mathbb{Y}_k$}
        \State $\mathbf{u}^a_{k-1}\gets\mathbf{u}^a_{k-1}/\overline{\mathbf{u}}^a_{k-1}$
        \State $\mathbf{v}_{k-1},\mathbf{u}_{k-1}\gets {\mathbf{u}_{k-1}(\overline{\mathbf{u}}_{k-1}^a,\mathbf{u}_{k-2}^a)}$

      \EndIf
    \doWhile{$C\hat{\mathbf{q}}_k^-\notin\mathbb{Y}_k$}

    \If{$\abs{y_k^m-C\hat{\mathbf{q}}_k^-}\leq \varepsilon$}
      \State $\hat{\mathbf{q}}_k\gets\hat{\mathbf{q}}_k^-$
    \Else 
      \State $P_k^-\gets AP_{k-1}A^T+Q$
      \State $L\gets P_k^-C^T/(CP_k^-C^T+R)$
      \State $\hat{\mathbf{q}}_k\gets \hat{\mathbf{q}}_k^-+L(y_k^m-C\hat{\mathbf{q}}_k^-)$
      $\mathbf{z}_{k-1}$
      \State $P_k\gets(I+LC)P_k^-$
    \EndIf
    \State $\mathbf{z}_k\gets A\mathbf{z}_{k-1}+B\mathbf{u}_{k-1}$
    \State $\mathbf{y}\gets\mathbf{y}\cup C\hat{\mathbf{q}}_k$
    \State $k\gets k + 1$
  \EndWhile % <---
  %\State $r\gets a\bmod b$
  %\State $\mathcal{M}_1\gets a\bmod b$
  %\While{$r\not=0$}\Comment{We have the answer if r is 0}
  %State $a\gets b$
  %\State $b\gets r$
  %\State $r\gets a\bmod b$
  %\EndWhile\label{euclidendwhile}
  %\State \textbf{return} $b$\Comment{The gcd is b}
  \EndProcedure
\end{algorithmic}

>>>>

The algorithm uses the concept to select the adjustments s.t. $\varphi_k(\mathbf{p}_k,\mathbf{c}_k^0)$ has the highest energy value. It guides the UAV to the new position $\mathbf{p}_{k+1}$ using the vector field of $\varPhi:=\varphi_k(\mathbf{p}_k,\mathbf{c}_k^0)$, deriving the direction to follow--the desired velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is the gradient, $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind $w\in\mathbb{R}$.

@@

Let us consider a Fourier series $h:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ for the purpose of energy consumption modeling of the mission
\begin{equation}\label{eq:fourier}
  h(t)=\sum_{n=0}^{r}{a_n\cos\left({nt/\xi}\right)+b_n\sin{\left({nt/\xi}\right)}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ the Fourier series coefficients.

@@

The energy cost of the computations is assessed using \powprof{}, an open-source modeling tool presented in our previous work~\cite{seewald2019coarse}, that measures software configurations empirically and builds a computational energy consumption model. Specifically, the tool builds a linear interpolation, one per each task. It requires the user to implement the mission as a ROS system with one or more ROS nodes changing the computational load by node-specific ROS parameters. A way to simulate the change of the computations.

@@

Given an initial stage $\mathcal{M}_0$ (TEE $\varphi_0$, tasks $\Psi$, computations $\mathbf{s}_0\in\mathbb{S}_0$, and adaptations $\mathbf{c}_0\in\mathbb{C}_0$), the algorithm produce a valid mission.

\begin{defn}[Valid mission]\label{def:valid}
  A mission is valid if for every stage $\mathcal{M}_{k-1}$ in $k\in[t_f]^+$ there exist a control sequence $\mathbf{u}_k^{a}$ which produce the stage $\mathcal{M}_k$
  \begin{equation}\begin{split}
    \mathbf{u}^a_{k}=\{(\mathbf{c}_{k},\mathbf{s}_{k})\mid&(\varphi(\mathbf{p}_{k-1},\mathbf{c}_{k-1}),\Psi(\mathbf{s_{k-1}}))\in\mathcal{M}_{k-1}\\
    &\Longrightarrow(\varphi(\mathbf{p}_{k},\mathbf{c}_{k}),\Psi(\mathbf{s}_{k}))\in\mathcal{M}_{k}\}.
  \end{split}\end{equation}
\end{defn}

Let us proof that if the mission is valid, the instantaneous energy consumption is a linear combination of the state from the Equation~(\ref{eq:state-perf}).

\begin{thm}[State output linearity]\label{thm:state-vs-energy}
  Consider the mission from Definition~\ref{def:mission}, the valid mission from~\ref{def:valid}, and assume Assumption~\ref{assm:periodic} holds. Likewise in Lemma~\ref{lem:state-vs-energy}, the model behaves ideally. 
  Then, the instantaneous energy consumption $y_k$ is a linear combination of the state $\mathbf{q}_k$ and $\mathcal{M}_k$ produce a valid mission $\mathcal{M}_{k+1}\,\forall k\in[t_f-1]$.
\end{thm}
\begin{proof}
  *\\
\end{proof}

@@

\begin{algorithmic}[1]
  \Procedure{Step}{$\mathcal{M}_{k-1}$, $\mathbf{q}_{k-1}$, $\mathbf{u}_{k-1}^a$, $\mathbf{u}_{k-2}^a$, $P_{k-1}$}\label{alg:step}

  \State $\mathbf{u}_{k-1}\gets {\mathbf{u}_{k-1}(\overline{\mathbf{u}}_{k-1}^a,\mathbf{u}_{k-2}^a)}$\label{alg:max_cont_sequence}

  \State $\mathbf{u}^0_{k-1}\gets\argmax_{\mathbf{u}}{\sum_{i=k-1}^{k+N-2}{l(\mathbf{q}_i,\mathbf{u}_i)+V_f(\mathbf{q}_{k+N-1})}}$\label{alg:mpc}

  \State $\hat{\mathbf{q}}_k^-\gets A\hat{\mathbf{q}}_{k-1} + B\mathbf{u}^0_{k-1}$
  \If{$C\hat{\mathbf{q}}_k^-\notin\mathbb{Y}_k$}\label{alg:output_constraints}
    \State $\mathbf{u}^a_{k-1}\gets\mathbf{u}^a_{k-1}/\{\overline{\mathbf{u}}^a_{k-1}\}$
    \State\Return{\Call{Step}{$\mathcal{M}_{k-1}$, $\mathbf{q}_{k-1}$, $\mathbf{u}^a_{k-1}$, $\mathbf{u}_{k-2}^a$,$P_{k-1}$}}
  \Else
    \If{$\abs{y_k^m-C\hat{\mathbf{q}}_k^-}\leq \varepsilon$}
      \State $\hat{\mathbf{q}}_k\gets\hat{\mathbf{q}}_k^-$\label{alg:evolution}
    \Else 
      \State $P_k^-\gets AP_{k-1}A^T+Q$\label{alg:kalman_start}
      \State $K\gets P_k^-C^T/(CP_k^-C^T+R)$
      \State $\hat{\mathbf{q}}_k\gets \hat{\mathbf{q}}_k^-+K(y_k^m-C\hat{\mathbf{q}}_k^-)$
      \State $P_k\gets(I+KC)P_k^-$\label{alg:kalman_end}
    \EndIf
    \State $\mathbf{u}_{k}^a\gets \mathbf{u}_{k-1}^a$
    \State\Return{$(\hat{\mathbf{q}}_k,\mathbf{u}_k^a)$}
  \EndIf
  %\State $r\gets a\bmod b$
  %\State $\mathcal{M}_1\gets a\bmod b$
  %\While{$r\not=0$}\Comment{We have the answer if r is 0}
  %State $a\gets b$
  %\State $b\gets r$
  %\State $r\gets a\bmod b$
  %\EndWhile\label{euclidendwhile}
  %\State \textbf{return} $b$\Comment{The gcd is b}
  \EndProcedure
  \Procedure{EADMPA}{$\mathcal{M}_0$, $\mathbf{p}_0$, $\mathbf{q}_0$}
  \State $k\gets 0$
%  \State $\mathbf{y}\gets \{\emptyset\}$
  \State $\mathbf{u}_{k-1}^a\gets\{\emptyset\}$
  \State $\mathbf{u}_{k}^a\gets \{(\mathbf{c}_0,\mathbf{s}_0)\mid(\varphi_0(\mathbf{p}_{0},\mathbf{c}_{0}),\Psi(\mathbf{s_{0}}))\in\mathcal{M}_{0}\}$
  \State $\mathcal{M}_k\gets\mathcal{M}_0$
  \State $\mathbf{p}_k\gets\mathbf{p}_0$
  \State $\mathbf{q}_k\gets\mathbf{q}_0$
  \While{$k \leq t_f$}

    \State $(\mathbf{q}_k,\mathbf{u}_k^a)\gets$\Call{Step}{$\mathbf{q}_{k}$, $\mathbf{u}^a_{k}$, $\mathbf{u}_{k-1}^a$}
    \State $\mathbf{p}_{k}=\mathbf{p}_k\dot{\mathbf{p}}_d(\mathbf{p}_k)/v$
    \State $\mathbf{u}_{k-1}^a\gets\mathbf{u}_{k}^a$

%    \State $\mathbf{y}\gets\mathbf{y}\cup C\hat{\mathbf{q}}_k$
    \State $k\gets k + 1$
  \EndWhile 
  \EndProcedure
\end{algorithmic}

%The algorithm per each time step updates the state. The position at line

Line~\ref{alg:max_cont_sequence} selects the maximum possible control from the control sequence at the time step $k-1$. Line~\ref{alg:mpc} uses robust output feedback model predictive control (MPC)~\cite{rawlings2017model} to select the optimal control $\mathbf{u}^0$ for a given horizon $N$ from the cost function
\begin{equation}\begin{split}
  l(\mathbf{q}_k,\mathbf{u}_k)&:=(1/2)(\mathbf{q}_k^TQ\mathbf{q}_k+\mathbf{u}_k^TR\mathbf{u}_k),\\
  V_f(\mathbf{q}_k)&:=(1/2)(\mathbf{q}_k^TP_f\mathbf{q}_k)
\end{split}\end{equation}
where matrices $Q\in\mathbb{R}^{j\times j},R\in\mathbb{R}^{l\times l}$ are all positive definite.

Follows a check if the mission can finish without the eventuality of the battery discharge (output constraints satisfaction) at line~\ref{alg:output_constraints}. The control sequence is then eventually updated.

Before the algorithm returns the energy evolution and the control sequence, a state estimator--the discrete-time Kalman filter~\cite{simon2006optimal} (lines~\ref{alg:kalman_start}--\ref{alg:kalman_end})--predicts the state $\mathbf{q}$ if the modeled instantaneous energy consumption diverging from the sensor's value $y_k^m\in\mathbb{R}$ more than a given $\varepsilon\in\mathbb{R}_{\geq  0}$.
